---
title: "NMR processing"
author: "Kaizad Patel"
date: 2020-01-31
output: github_document
---
Last updated: `r Sys.Date()`  


This script is to batch-analyze processed NMR spectra (phase corrected, baseline corrected, deconvoluted, peak picked).  

Select the appropriate binning option and then plot and calculate relative abundances.  


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE,message=FALSE,warning=FALSE,
                      collapse = TRUE,
                      comment = "#>" #, fig.path = "images/"
                      )

source("1-processing.R")
```
----
Example spectra:  
spectra processed in MestreNova, with automated Global Spectral Deconvolution.   
Water and solvent (DMSO) regions removed for graphing purposes. Still needs some cleaning. 

```{r, fig.width=12, fig.height=6}
spectra2 = 
  spectra2 %>% 
  filter(!(ppm>DMSO_start&ppm<WATER_stop))
gg_nmr+
  geom_path(data=spectra2, aes(x = ppm, y = intensity, color = source))+
# if you need to add labels:
  annotate("text", label = "aliphatic", x = 1.4, y = -500)+
  annotate("text", label = "O-alkyl", x = 3.5, y = -500)+
  annotate("text", label = "alpha-H\n (protein)", x = 4.45, y = -600)+
  annotate("text", label = "aromatic", x = 7, y = -500)+
  annotate("text", label = "amide", x = 8.1, y = -500)+
  annotate("text", label = "\n\nWATER", x = 3.5, y = Inf)+
  annotate("text", label = "\n\nDMSO", x = 2.48, y = Inf)+
  geom_vline(xintercept = WATER_start, linetype="longdash")+
  geom_vline(xintercept = WATER_stop, linetype="longdash")+
  geom_vline(xintercept = DMSO_start, linetype="dashed")+
  geom_vline(xintercept = DMSO_stop, linetype="dashed")
```

Including solvents 

```{r, fig.width=12, fig.height=6}
gg_nmr+
  geom_path(data=spectra[(spectra$ppm>0&spectra$ppm<10),], aes(x = ppm, y = intensity, color = source))+
  ylim(-700,25000)+
  #expand_limits(y=c(NA, 35000))+
  # if you need to add labels:
  annotate("text", label = "aliphatic", x = 1.4, y = -500)+
  annotate("text", label = "O-alkyl", x = 3.5, y = -500)+
  annotate("text", label = "alpha-H\n (protein)", x = 4.45, y = -600)+
  annotate("text", label = "aromatic", x = 7, y = -500)+
  annotate("text", label = "amide", x = 8.1, y = -500)+
  annotate("text", label = "\n\nWATER", x = 3.5, y = Inf)+
  annotate("text", label = "\n\nDMSO", x = 2.48, y = Inf)+
  geom_vline(xintercept = WATER_start, linetype="longdash")+
  geom_vline(xintercept = WATER_stop, linetype="longdash")+
  geom_vline(xintercept = DMSO_start, linetype="dashed")+
  geom_vline(xintercept = DMSO_stop, linetype="dashed")
```
