---
title: "nmrrr - general workflow"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

This vignette represents the general workflow for processing NMR data using the *nmrrr* package. 

This package works only on processed data (e.g., phase-corrected, baseline-corrected, peak-picked, etc.). 
For tips on processing NMR data in MestreNova, check out the [repository wiki](https://github.com/kaizadp/nmrrr/wiki).

---

<details>
  <summary>A note on the dataset - click to open</summary>
  
The data used in this example are a subset of the dataset generated for Patel et al. "Soil carbon dynamics during drying vs. rewetting: Importance of antecedent moisture conditions". [DOI: 10.1016/j.soilbio.2021.108165](https://doi.org/10.1016/j.soilbio.2021.108165). 

Soils were subjected to two treatments:

a. drought: soils were dried and then rewet to 50 % moisture (w/w)
b. flood: soils were saturated and then dried to 50 % moisture (w/w)
  
</details>


---

Notes:

- If you have individual files for your multi-sample batch, save all spectra files in one folder and peaks files in another folder
- All input files must be in .csv format

---

## PART I: Setup (parameters)

```{r}
library(tidyverse)
library(readxl)
```

### Step 1. `source` the functions 

(*once we have the package, we won't have to `source`, just load the package - KP* ) 


```{r}
source("code/nmrrr_workflow/1-functions_processing.R")
source("code/nmrrr_workflow/3-functions_graphs.R")
```

### Step 2. set the bins using BINSET = ...

```{r}
BINSET = "Clemente2012"
bins_dat = set_bins(BINSET)
```

### Step 3. Set input directories

```{r}
SPECTRA_FILES = "data/KFP_hysteresis/spectra/"
PEAKS_FILES = "data/KFP_hysteresis/peaks/"
```

---

## PART II: Processing spectra files


```{r}
spectra_processed = import_nmr_spectra_data(SPECTRA_FILES)
```

This function will:

- import all the files from the SPECTRA_FILES location, 
- combine them into a single dataframe,
- clean the dataframe,
- and generate a dataframe with columns for ppm-shift, intensity, and sample-ID


```{r}
head(spectra_processed)
```


---

## PART III: Plotting the spectra

The `gg_spectra()` function will plot all the spectra present in the `spectra_processed` file.

```{r}
gg_spectra(dat = spectra_processed, 
           LABEL_POSITION = 3, 
           aes(x = ppm, y = intensity, group = source, color = source),
           STAGGER = 0.5) + 
  labs(subtitle = paste("binset:", BINSET))+
  ylim(0, 3.5)
```

The compound classes (bins) are labelled in the graph. Adjust the position (y-axis) using `LABEL_POSITION = ...`.

Overlaid spectra can be staggered along the y-axis using `STAGGER - ...`.

This function makes use of `ggplot2` capabilities, and the plot can therefore be customized using `ggplot2` nomenclature.


<details>
  <summary>Additional examples using `gg_spectra()`  - *click to open*</summary>

```{r}
gg_spectra(dat = spectra_processed, 
           LABEL_POSITION = 3, 
           aes(x = ppm, y = intensity, group = source),
           STAGGER = 0.5) + 
  labs(subtitle = paste("binset:", BINSET))+
  ylim(0, 3.5)
```

</details>


---


## PART IV: Processing peaks data

When exporting peaks data from MestreNova, the data are generally split across multiple columns.

The `import_nmr_peaks()` function will import the files in that format and then collapse all the columns for a single long-form dataframe.

```{r}
peaks_processed = import_nmr_peaks(PEAKS_FILES)
```

```{r}
head(peaks_processed)
```


**Note: The user may want to assign additional filtering steps to filter certain flagged data points, e.g. impurities, weak peaks, etc. ** 

For this current dataset, because of the strong influence of water peaks in the o-alkyl region, we exclude that region from our calculations.

```{r}
peaks_processed2 = 
  peaks_processed %>% 
  filter(group != "oalkyl")
```


---

## PART V: Exporting the processed data

The processed dataframes generated here can be exported using the usual R export functions, e.g. `write.csv()`, `write.table()`, etc.  


---

<details>
  <summary>Session Info - click to open</summary>

Date run: `r Sys.Date()`


```{r, echo=FALSE}
sessionInfo()
```

</details>

---